\newpage
\begin{center}
%	\noindent
	\LARGE
	\dissertationtitle \\
	\vspace*{\stretch{1}}
\end{center}

{\Huge \bf Abstract}
\vspace{24pt} 

Computational biology often involves working with high-dimensional data. Penalized regression methods are often used on such data, as they can perform feature selection effectively. Several approaches for network-constrained regression have been suggested in literature over the recent years. They use prior knowledge in the form of a network to exploit known relationships between predictors. Parameter tuning for the various implemented methods is performed on synthetic datasets. 

An approach for cooperative parameter tuning in the context of multiple alternative methods that share common input and goals is suggested. The aim is to simultaneously tune the different regression methods iteratively, in a way that increases agreement between their coefficient estimates. Neighboring values on the tuning parameter grid are considered for each method and iteration, selecting the set of values that achieves the largest correlation with the averaged coefficients of all other methods for the previous iteration. This process converges, given enough iterations and granularity of the tuning grids. Several possible modifications to this approach are suggested and discussed.

We also suggest a simple approach to aggregate the coefficients produced by the various regression methods. Each predictor is considered relevant if it corresponds to a non-zero coefficient in a certain fraction of the underlying methods. Once a consensus regarding predictor relevance is reached through this form of voting, ordinary least squares estimation is used to fit only the relevant predictors to the data.

The traditional approach of hyperparameter tuning by minimization of the prediction mean squared error is implemented alongside our suggested approach. Comparison of both tuning approaches is performed on synthetic datasets. Analysis of similarities between the various regression methods is carried out. An optimal set of tuning parameters is assembled for each regression method for use on real epigenetic breast cancer data. Gene methylation and expression data has been processed with the implemented algorithms. We explore how the expression level of each gene is affected by the methylation levels of related genes.

\newpage
Total word count: \wordcount

* As calculated by TeXcount [ https://www.ctan.org/pkg/texcount ]
\vspace*{\fill}
