\chapter{Introduction}
\pagenumbering{arabic} 
\setcounter{page}{1}
Epigenetics \cite{holliday2006epigenetics} studies the heritable traits that cannot be explained by changes in the DNA sequence. Examples of epigenetic mechanisms are DNA methylation and histone modification. These mechanisms adjust the expression level of genes \cite{jaenisch2003epigenetic}, which allows organisms to dynamically adapt to changes in the environment.

Disruption of gene expression levels is related to the development of various diseases \cite{egger2004epigenetics}. For example, the epigenetic deactivation of certain tumor suppressor genes commonly leads to the development of cancer \cite{esteller2008epigenetics}. The expression levels of certain genes can therefore be used as additional tools in early diagnostics of cancer, as prognosis factors and as predictors of response to treatment.

Understanding of the relationship between DNA methylation and gene expression is important for both cancer prevention and epigenetic disease treatment. We have used gene methylation and expression level data from breast cancer patients, obtained from The Cancer Genome Atlas (TCGA) \cite{cancer2012comprehensive}, to explore this relationship. Among the goals of this project is to produce a map that shows the methylation of which genes affects the expression levels of each gene.

Several linear regression methods \cite{tibshirani1996regression,zou2005regularization,li2008network,li2010variable,pan2010incorporating,luo2012two,kim2013network} have been implemented and considered for use with the real breast cancer data. The hyperparameters for each method have been tuned with the use of synthetic datasets as suggested in \cite{li2008network}. This is done because ground truth remains unknown for the relationship between gene methylation and expression.

The various regression methods discussed in this project minimize different cost functions. As a result, each method exhibits specific strengths and weaknesses. The relative performance of the methods depends on the dataset used. For this reason it is impossible to predict their effectiveness on an arbitrary real data set with no access to ground truth. We have developed a simple approach to merge the results of coefficient estimation produced by the different regression methods. It is designed to balance the behavior of any individual method. Each of the predictor variables is considered important if it has non-zero coefficients in a fraction of the methods above a given threshold. This approach for variable selection in practice implements a voting system where each predictor must achieve a certain electoral threshold to be selected. Ordinary least squares estimation is then performed only using the set of selected variables.

A novel method of hyperparameter tuning is developed as an alternative to the widely used method of minimizing the cross-validated mean squared test error. In our context we have a collection of regression methods that operate on the same training data and share a common goal - to correctly identify the relationship between predictor and target variables. Instead of tuning the various regression methods independently, our approach performs cooperative hyperparameter tuning on all methods simultaneously. It uses an iterative algorithm to increase the similarity of estimated coefficients for the various methods by tuning their hyperparameters. For each method and iteration, the method's parameter grid neighborhood is searched for a set of parameters that maximizes the correlation between its estimated coefficients and the averaged estimates of all other methods for the previous iteration. When this process converges a set of parameters is defined for each method that maximizes the overall agreement across the whole set of methods. Several possible modifications to this approach are suggested and discussed.

The following chapters discuss the theoretical and implementation details of each linear regression approach. Afterwards, we introduce the synthetic dataset generation process used in our simulations. We then define our estimate merging procedure and our proposed cooperative hyperparameter tuning approach, along with its possible modifications. In the simulations carried out we compare the various regression approaches, as well as the traditional and our proposed cooperative parameter tuning approach. The similarities between estimates of the various regression methods have been explored. We also look into the effect that a choice of hyperparameter tuning procedure has on these similarities. The final chapter discusses experiments on a real breast cancer dataset performed in order to explore how the expression level of each gene is affected by the methylation levels of related genes.
